---
title: "r4ds applications"
author: "Gregory Park"
date: "2026-02-27"
output: github_document
---

## Loading
I will be working with "gapminder" dataset, which features country, continent, year, life expectancy, population, and gdp per capita
```{r, message=FALSE}
library(tidyverse); library(gapminder)

view(gapminder)
gapminder 
?gapminder
```

## Row work
"filter" if you want specific rows based on column rules
```{r}
gapminder |> 
  filter(year > 1990)

gapminder |> 
  filter(continent == "Asia" | continent == "Europe")

# combines | and & functions
gapminder |> 
  filter(country %in% c("Afghanistan","Albania"))

# dplyr functions never modify their inputs
new <- gapminder |> 
  filter(year == 2007)
view(new)
```
"arrange" changes the order of rows
```{r}
gapminder |> 
  arrange(desc(year))
```
"distinct" finds unique rows
```{r}
gapminder |> 
  distinct(continent)

gapminder |> 
  distinct(continent, year, .keep_all = TRUE)
```
And "count" will provide row counts
```{r}
gapminder |> 
  count(continent, sort = TRUE)
```

## Columns
"mutate" adds columns based on existing ones
```{r}
gapminder |> 
  mutate(pop_mil = pop / 1e6,
         .after = "pop"
         )
```
"select" will provide a subset of columns
```{r}
gapminder |> 
  select(!year:pop)

gapminder |> 
  mutate(pop_mil = pop / 1e6) |> 
  select(contains("pop"))
```
"rename" put the new name on the left
```{r}
gapminder |> 
  rename(population = pop)
```

