---
title: "Data Tidying"
output: github_document
---

```{r, message=FALSE}
library(tidyverse); library(tidyr)
```

There are three interrelated rules that make a dataset tidy:\
  1. Each variable is a column; each column is a variable.\
  2. Each observation is a row; each row is an observation.\
  3. Each value is a cell; each cell is a single value.

We may need to pivot data, with variables in columns and observations in rows.
```{r}
data("billboard")
billboard
```
Here we have the per-week rankings in rows. Want to create a separate column for rank.
```{r}
new_bill <- billboard |> 
  pivot_longer(
    cols = starts_with("wk"),
    names_to = "week",
    values_to = "rank",
    values_drop_na = TRUE # some songs didn't chart every week
  ) |> 
  mutate(
    week = parse_number(week) # extracts first number from string
  )
new_bill
```
Plotting to visualize how long tracks stay on chart
```{r}
new_bill |> 
  ggplot(aes(x = week, y = rank, group = track)) +
  geom_line(alpha = .25) +
  scale_y_reverse()
```

